---
title: "SOCIOL 722 - Week 1 Assignment"
format: html

execute: 
  echo: true
  warning: false
  message: false
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(gapminder)
library(tinytable)
library(knitr)
library(kableExtra)
library(lubridate)

theme_set(theme_light())

here::here()

data <- read_csv("books.csv")

```

# Dataset

The dataset is from Goodreads, a book review website. The source is: https://www.kaggle.com/datasets/jealousleopard/goodreadsbooks

# Univariate Plots

## Density Plot

Density plot of average book ratings.

```{r}
ggplot(data,
       aes(x = average_rating)) +
    geom_density()
```

## Histogram

Histogram of average book ratings.

```{r}
                 
ggplot(data,
       aes(x = average_rating)) +
  geom_histogram(binwidth = 1,
                 boundary = 0,
                 color = "white")

```

## Bar Graph

Bar graph of publication year.

```{r}

past_20_years <- data |> 
  mutate(publication_year = year(mdy(publication_date))) |> 
  filter(publication_year > 2004)


ggplot(past_20_years,
       aes(x = publication_year)) +
  geom_bar()

```

# Bivariate Plots

## Scatterplot

Scatterplot for number of pages and average rating of top 100 books (by rating count).

```{r}
top_100_books <- data |> 
  top_n(n=100, wt = ratings_count) 
  

ggplot(top_100_books,
       aes(x = num_pages,
           y = average_rating)) +
  geom_point()
```

## Bar Graph

Bar graph of book count for the top 10 publishers.

```{r}
top_10_publishers <- data |> 
  group_by(publisher) |> 
  tally() |> 
  rename("book_count" = "n") |> 
  top_n(n = 10, wt = book_count)


ggplot(top_10_publishers,
       aes(x = publisher,
           y = book_count)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(labels = function(publisher) str_wrap(publisher, width = 10))
```

## Strip Plot

Strip plot showing average rating for top 100 books published in the past 20 years.

```{r}

data |>
  mutate(publication_year = year(mdy(publication_date))) |> 
  filter(publication_year > 2004) |>
  top_n(n = 100, wt = ratings_count) |> 
  ggplot(aes(x = publication_year,
           y = average_rating)) +
  geom_jitter(height = .1,
              width = .1, 
              alpha = .2)

```

## Time Plot (Line)

Time plot showing number of cumulative books per year for the top 10 publishers.

```{r}

data_with_publication_year <- data |> 
  mutate(publication_year = year(mdy(publication_date)))

top_10_publisher_books <- data_with_publication_year |> 
  group_by(publisher) |> 
  arrange(mdy(publication_date)) |> 
  mutate(id = row_number()) |> 
  filter(publisher %in% top_10_publishers$publisher)

top_10_publisher_books |> 
  rename("book_count" = "id") |> 
  ggplot(aes(x = publication_year,
             y = book_count,
             color = publisher)) +
  geom_line()

```
