---
title: "Week 10 - Homework"
format: 
  html:
    embed-resources: true
editor: visual

author: "David Gonzalez Chavez"

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(gssr)


theme_set(theme_light())

```

# Data

The data for this homework assignment comes from the 2022 GSS.

```{r}

gss2022 <- gss_get_yr(year = 2022) |> # get 2022 GSS data
  haven::zap_labels()                 # remove labels

```

# Analysis #1: Degree Attainment by Sex

The first analysis will be between men and women. The analysis will be differences in their church attendance.

```{r}

sex_church <- gss2022 |> 
  select(sex, attend) |> # select the sex variable and church attendance (comparison) variables
  drop_na() |> 
  mutate(sex_label = if_else(sex == 1, "Male", "Female")) |> 
  mutate(sex = if_else(sex == 1, 1, 0)) |> 
  mutate(attend_label = if_else(attend==0, "Does Not Attend Church", "Attends Church")) |> # 0 is never, 1+ is attends 
  mutate(attend = if_else(attend == 0, 1, 0)) # if attend is 0, they do not attend church

sex_church_plot <- sex_church |>
  group_by(sex_label, attend_label) |> # group by sex and church attendance
  summarize(n = n()) |> # get count with that combo
  mutate(percent = n/sum(n)) # get percent of group for each subgroup

ggplot(sex_church_plot,
       aes(x = attend_label,
           fill = sex_label,
           y = percent)) +
  geom_bar(position = "dodge",
           stat = "identity") +
  labs(x = "Church Attendance",
       fill = "Sex") +
  scale_y_continuous(labels = scales::percent)

```

## Difference in Probability Regression

```{r}
mod1_sex_church <- glm(attend ~ sex,
            data = sex_church,
            family = binomial(link = "identity"))

summary(mod1_sex_church)
```

The difference in probabilities regression suggests that men are `{r} abs(round(summary(mod1_sex_church)$coefficients[2],4))*100` percentage points more likely to not attend church than women. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in probability).

## Risk Ratio Regression

```{r}

mod2_sex_church <- glm(attend ~ sex,
            data = sex_church,
            family = binomial(link = "log"))

summary(mod2_sex_church)

exp(summary(mod2_sex_church)$coefficients[2])

```

The risk ratio regression suggests that men are `{r} (exp(summary(mod2_sex_church)$coefficients[2])*100) - 100` percent more likely to not attend church than women. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in risk).

## Logistic Regression

```{r}

mod3_sex_church <- glm(attend ~ sex,
            data = sex_church,
            family = binomial(link = "logit"))

summary(mod3_sex_church)

exp(summary(mod3_sex_church)$coefficients[2])

```
The logistic regression suggests that the odds of men not attending church are `{r} exp(summary(mod3_sex_church)$coefficients[2])` times higher than for women. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in odds).

# Analysis #2: X-Rated Movies by Sex

The second analysis will be between men and women. The analysis will be differences in whether they watched an X-rated movie in the last year. 

```{r}

sex_xrated <- gss2022 |> 
  select(sex, xmovie) |> # select the sex variable and x-rated movie (comparison) variables
  drop_na() |> 
  mutate(sex_label = if_else(sex == 1, "Male", "Female")) |> 
  mutate(sex = if_else(sex == 1, 1, 0)) |> 
  mutate(xmovie_label = if_else(xmovie==1, "Did Watch", "Did Not Watch")) |> # 1 is did watch, 2 is no
  mutate(xmovie = if_else(xmovie == 1, 1, 0)) # if xmovie is 0, they do not consume x-rated movie

sex_xrated_plot <- sex_xrated |>
  group_by(sex_label, xmovie_label) |> # group by sex and xrated movie watching
  summarize(n = n()) |> # get count with that combo
  mutate(percent = n/sum(n)) # get percent of group for each subgroup

ggplot(sex_xrated_plot,
       aes(x = xmovie_label,
           fill = sex_label,
           y = percent)) +
  geom_bar(position = "dodge",
           stat = "identity") +
  labs(x = "X-Rated Movie in Past Year",
       fill = "Sex") +
  scale_y_continuous(labels = scales::percent)

```

## Difference in Probability Regression

```{r}
mod1_sex_xrated <- glm(xmovie ~ sex,
            data = sex_xrated,
            family = binomial(link = "identity"))

summary(mod1_sex_xrated)
```

The difference in probabilities regression suggests that men are `{r} abs(round(summary(mod1_sex_xrated)$coefficients[2],4))*100` percentage points more likely to have watched an x-rated movie in the past year than women. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in probability).

## Risk Ratio Regression

```{r}

mod2_sex_xrated <- glm(xmovie ~ sex,
            data = sex_xrated,
            family = binomial(link = "log"))

summary(mod2_sex_xrated)

exp(summary(mod2_sex_xrated)$coefficients[2])

```

The risk ratio regression suggests that men are `{r} (exp(summary(mod2_sex_xrated)$coefficients[2])*100) - 100` percent more likely to have watched an x-rated movie in the past year than women. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in risk).

## Logistic Regression

```{r}

mod3_sex_xrated <- glm(xmovie ~ sex,
            data = sex_xrated,
            family = binomial(link = "logit"))

summary(mod3_sex_xrated)

exp(summary(mod3_sex_xrated)$coefficients[2])

```
The logistic regression suggests that the odds of men having watched an x-rated movie in the past year are `{r} exp(summary(mod3_sex_xrated)$coefficients[2])` times higher than for women. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in odds).

# Analysis #3: Happiness by Marital Status

The third analysis will be between married and unmarried respondents. The analysis will be differences in whether they reported being very happy. 

```{r}

marriage_happiness <- gss2022 |> 
  select(marital, happy) |> # select the marriage variable and x-rated movie (comparison) variables
  drop_na() |> 
  mutate(marriage_label = if_else(marital == 1, "Actively Married", "Not Currently Married")) |> #1 = married, 2+ is unmarried, widowed, etc. 
  mutate(marriage = if_else(marital == 1, 1, 0)) |> 
  mutate(happy_label = if_else(happy==1, "Very Happy", "Not Very Happy")) |> # 1 is very happy, 2 pretty happy, 3 not too happy
  mutate(happy = if_else(happy == 1, 1, 0)) # if happy is 1, very happy

marriage_happiness_plot <- marriage_happiness |>
  group_by(marriage_label, happy_label) |> # group by marriage and happiness
  summarize(n = n()) |> # get count with that combo
  mutate(percent = n/sum(n)) # get percent of group for each subgroup

ggplot(marriage_happiness_plot,
       aes(x = happy_label,
           fill = marriage_label,
           y = percent)) +
  geom_bar(position = "dodge",
           stat = "identity") +
  labs(x = "Happiness",
       fill = "Marital Status") +
  scale_y_continuous(labels = scales::percent)

```

## Difference in Probability Regression

```{r}
mod1_marriage_happiness <- glm(happy ~ marriage,
            data = marriage_happiness,
            family = binomial(link = "identity"))

summary(mod1_marriage_happiness)
```

The difference in probabilities regression suggests that married people are `{r} abs(round(summary(mod1_marriage_happiness)$coefficients[2],4))*100` percentage points more likely to be very happy than unmarried people. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in probability).

## Risk Ratio Regression

```{r}

mod2_marriage_happiness <- glm(happy ~ marriage,
            data = marriage_happiness,
            family = binomial(link = "log"))

summary(mod2_marriage_happiness)

exp(summary(mod2_marriage_happiness)$coefficients[2])

```

The risk ratio regression suggests that married people are `{r} exp(summary(mod2_marriage_happiness)$coefficients[2])*100 - 100` percent more likely to be very happy than unmarried people. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in risk).

## Logistic Regression

```{r}

mod3_marriage_happiness <- glm(happy ~ marriage,
            data = marriage_happiness,
            family = binomial(link = "logit"))

summary(mod3_marriage_happiness)

exp(summary(mod3_marriage_happiness)$coefficients[2])

```
The logistic regression suggests that the odds of married people being very happy are `{r} exp(summary(mod3_marriage_happiness)$coefficients[2])` times higher than for unmarried people. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in odds).

# Analysis #4: Financial Satisfaction by Marital Status

The fourth analysis will be between married and unmarried respondents. The analysis will be differences in financial satisfaction. 

```{r}

marriage_finsatisfaction <- gss2022 |> 
  select(marital, satfin) |> # select the marriage variable and x-rated movie (comparison) variables
  drop_na() |> 
  mutate(marriage_label = if_else(marital == 1, "Actively Married", "Not Currently Married")) |> #1 = married, 2+ is unmarried, widowed, etc. 
  mutate(marriage = if_else(marital == 1, 1, 0)) |> 
  mutate(satfin_label = if_else(satfin==1, "Well Satisfied", "Not Well Satisfied")) |> # 1 is pretty well satisfied, 2+ is more/less and not satisfied
  mutate(satfin = if_else(satfin == 1, 1, 0)) # if satfin is 1, well satisfied

marriage_finsatisfaction_plot <- marriage_finsatisfaction |>
  group_by(marriage_label, satfin_label) |> # group by marriage and satisfaction
  summarize(n = n()) |> # get count with that combo
  mutate(percent = n/sum(n)) # get percent of group for each subgroup

ggplot(marriage_finsatisfaction_plot,
       aes(x = satfin_label,
           fill = marriage_label,
           y = percent)) +
  geom_bar(position = "dodge",
           stat = "identity") +
  labs(x = "Satisfaction With Current FInancial Situation",
       fill = "Marital Status") +
  scale_y_continuous(labels = scales::percent)

```

## Difference in Probability Regression

```{r}
mod1_marriage_finsatisfaction <- glm(satfin ~ marriage,
            data = marriage_finsatisfaction,
            family = binomial(link = "identity"))

summary(mod1_marriage_finsatisfaction)
```

The difference in probabilities regression suggests that married people are `{r} abs(round(summary(mod1_marriage_finsatisfaction)$coefficients[2],4))*100` percentage points more likely to be satisfied with their financial situation than unmarried people. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in probability).

## Risk Ratio Regression

```{r}

mod2_marriage_finsatisfaction <- glm(satfin ~ marriage,
            data = marriage_finsatisfaction,
            family = binomial(link = "log"))

summary(mod2_marriage_finsatisfaction)

exp(summary(mod2_marriage_finsatisfaction)$coefficients[2])

```

The risk ratio regression suggests that married people are `{r} exp(summary(mod2_marriage_finsatisfaction)$coefficients[2])*100 - 100` percent more likely to be satisfied with their financial situation than unmarried people. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in risk).

## Logistic Regression

```{r}

mod3_marriage_finsatisfaction <- glm(satfin ~ marriage,
            data = marriage_finsatisfaction,
            family = binomial(link = "logit"))

summary(mod3_marriage_finsatisfaction)

exp(summary(mod3_marriage_finsatisfaction)$coefficients[2])

```
The logistic regression suggests that the odds of married people being satisfied with their financial situation are `{r} exp(summary(mod3_marriage_finsatisfaction)$coefficients[2])` times higher than for unmarried people. The p-value for this estimate is less than our alpha level of .05, therefore we can reject the null hypothesis (no difference in odds).
