---
title: "Week 2 - Homework"
format: html

author: "David Gonzalez Chavez"

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(tinytable)

theme_set(theme_light())

```

# Creating the population.

The following code creates the working population: an imaginary town where 150K people live. Of the 150k people, 120k prefer dogs to cats. This represents a population parameter of 80% for preferring dogs.

```{r}

population <- tibble(id = 1:150000) |>  # initialize with 150K rows
  mutate(prefers_dogs = if_else(id <= 120000, 1L, 0L))  # first 120k individuals prefer dogs (1)

population |> 
  summarize(mean = mean(prefers_dogs)) # sanity check to confirm that 80% of people prefer dogs

```

# Drawing the sample.

The following code first draws 3 sets of 2500 samples of varying sample sizes from the population: 500, 1000, and 2500. It then draws a plot comparing the sample means of each different group. The result is a density plot demonstrating that each increasing sample size results in a tighter concentration of sample means, indicating a smaller IQR (and therefore greater precision).

```{r}

set.seed(12345)

create_sample <- function(samp_size = 1000, samp_quantity = 2500) { # default n = 1000
  tibble(1:samp_quantity) |> # create a tibble with all the samples
    rowwise() |> # treat each row independently
    mutate(sample_mean = (
             slice_sample(population, n = samp_size, replace = FALSE) |> # take a sample from the population with the sample size being provided in the function call
               summarize(sample_mean = mean(prefers_dogs)) |>  # take an average of prefers_dogs
               as.double()
    )) |> 
    mutate(sample_size_group = paste(samp_size, " Sample Size")) # label the sample size group
}

samples_500 <- create_sample(samp_size = 500, samp_quantity = 2500) # create a tibble with 2500 samples of size 500 each
samples_1000 <- create_sample(samp_size = 1000, samp_quantity = 2500) # create a tibble with 2500 samples of size 1000 each
samples_2500 <- create_sample(samp_size = 2500, samp_quantity = 2500) # create a tibble with 2500 samples of size 2500 each


combined_df <- bind_rows(samples_500, samples_1000, samples_2500) # combine all tibbles to create a plot

ggplot(combined_df,
       aes(x = sample_mean, # the X is individual sample means (i.e., 2500 different points per sample size group)
              color = sample_size_group)) + # identify each by their sample size group label
  geom_density() # draw a density plot


```
